---
title: "Power B analysis"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(pwr)        
library(emmeans)


files <- c(
  c(
    "0j17p08k.csv",
    "1dg2qhj2b3.csv",
    "1e17ytgq.csv",
    "1fqrd6qr.csv",
    "1lmu0rmy.csv",
    "1orqqvyp.csv",
    "1qr6nnt1.csv",
    "1sftbbwm.csv",
    "2bomnjjb.csv",
    "2ea7uxyc.csv",
    "2ewef3gm.csv",
    "2odgntx7.csv",
    "3e3v39z8.csv",
    "3jtr11xk.csv",
    "3og97j2k.csv",
    "3ut1qsdd.csv",
    "4gjn1seg.csv",
    "4qcemy4k.csv",
    "4yxop4r3.csv",
    "5lqc6nvcre.csv",
    "5vrdpq0t.csv",
    "7f2wtp1j.csv",
    "7lhlpfus.csv",
    "7mza4pym.csv",
    "9kd8a7eb.csv",
    "9nvsempv62.csv",
    "20frs6ol.csv",
    "31by1doq.csv",
    "91mfkpae.csv",
    "120ypmwg.csv",
    "352jd0sb.csv",
    "970gbe0a.csv",
    "a7afpb9w.csv",
    "adbunmf0.csv",
    "ajl3gqra.csv",
    "bn7akg4o.csv",
    "bo3a253d.csv",
    "bqlvmxzf.csv",
    "c1m9yfql.csv",
    "cktudb9s.csv",
    "dsfc282a.csv",
    "e34duc56.csv",
    "el2hvftl.csv",
    "gfzoatpu.csv",
    "h8nlnku2.csv",
    "h6253x9m.csv",
    "hb3mv8ph.csv",
    "hb4brkxq.csv",
    "hs2xklzn.csv",
    "hwsqhjot.csv",
    "jrj7xzuf.csv",
    "k2b5yzkl.csv",
    "k8rnkquz.csv",
    "kl8k8j1y.csv",
    "lta397lq.csv",
    "luzwjk24.csv",
    "lwvpwcq0.csv",
    "m1eabfx1.csv",
    "n8adea4h.csv",
    "na6hh0ua.csv",
    "nflshhgq.csv",
    "o92u56uh.csv",
    "ogyjt8yg.csv",
    "ojv4jxxa.csv",
    "olol1tte.csv",
    "orybxbxu.csv",
    "p3taub1f.csv",
    "pco9vqna.csv",
    "pm5zycnt.csv",
    "ppt63ukd.csv",
    "q330tno2.csv",
    "qbwjdear.csv",
    "qfm8ks37.csv",
    "qgzx1bfu.csv",
    "r0pm70sc.csv",
    "r1o8o6o4.csv",
    "repvu995.csv",
    "rxur173z.csv",
    "s3j977x7.csv",
    "szh4zjou.csv",
    "tf6nb9bm.csv",
    "tpt94g2b.csv",
    "tw3k60gy.csv",
    "txdtrb48.csv",
    "u0pme1hk.csv",
    "u1s85lkk.csv",
    "u2fs8oqd.csv",
    "ucyeq6mc.csv",
    "uu3k1546.csv",
    "vg74l3ea.csv",
    "vgp86ou0.csv",
    "vp4hn41a.csv",
    "w3yqa2f3.csv",
    "xthmgyn8.csv",
    "y14x4t45.csv",
    "yu3vjzn3.csv",
    "z25ltc96.csv"
  )
)

# Read and combine all files from ../stims
data.raw <- files %>%
  set_names() %>%
  map_dfr(
    ~ suppressMessages(
        suppressWarnings(
          read_csv(
            file.path("../data/pilotb", .x),
            show_col_types = FALSE
          ) %>%
          mutate(rt = as.numeric(rt))
        )
      ),
    .id = "source_file"
  )

stims.raw <- read.csv("../stims/stims_abridged_all_info.csv")

# Remove filler and practice entries
# Focus on only target word for now (i.e. deepest NP in clause)
data <- data.raw %>%
  select(!stimulus) %>%
  left_join(stims.raw %>% select(item_id, target_word, stimulus), multiple = "first", by = join_by(item_id)) %>%
  mutate(rt = as.numeric(rt),
         log_rt = log(rt),
         whword = factor(ifelse(condition == 'a', 'object who', 
                         ifelse(condition == 'b', 'how',
                                ifelse(condition == 'c', 'that',
                                       'why'))), levels = c("object who", "how", "that", "why"))) %>% # condition = wh-word used in sentence
  filter(item_type != 'filler' & item_type != 'practice' & too_fast == FALSE) %>%
  filter(whword != "object who" | word_index < 19) %>%
  mutate(is_target_word = (target_word == word)) # select only deepest embedded NP
  

write.csv(data, "../data/pilot_b_data.csv")
```

Here we only include reading time observations taken at the most deeply embedded N. Further, we eliminate all observations where the reading time \< 100 ms (done via the condition too_fast == FALSE).

```{r}
data_embedded <- data %>% filter(is_target_word) # select only depeest embedded NP
```

Description of each field:

-   source_file: the original file in the directory 'data' storing a given participant's data

-   participant_id: 4-char participant id provided by Prolific

-   rt: reading time (log_rt: log transformed reading time)

-   condition: value from 'a'-'d' describing type of stimulus sentence

-   word: word at which reading time was measured

-   whword: one of four values "object who", "why", "how", or "that" – denotes the type of relative clause in the stimulus

```{r}
summary(data_embedded)
```

```{r}
# Compute mean reaction time per condition
agr <- data_embedded %>%
  group_by(whword) %>%
  summarise(n = n(), se = sd(log_rt) / sqrt(n()), y = mean(log_rt)) %>%
  mutate(
    whword = factor(
      whword,
      levels = c("object who", "how", "that", "why")
    ) )# Order whword elements to be object who, how, that, why

agr


ggplot(agr, aes(x = whword, y = y, fill = whword)) +
  geom_bar(stat = "identity", width = 0.6, color = "black", linewidth = 1) +
  geom_errorbar(aes(ymin = y - se, ymax = y + se ),   
                width = 0.08,   linewidth = 1 ) +
  scale_fill_manual(values = c("object who" = "blue",
                               "how" = "darkred",
                               "that" = "green",
                               "why" = "purple")) +
  coord_cartesian(ylim = c(5, 5.6)) +
  labs(x = "wh-types", y = "Reaction time (log ms)") +
  ggtitle("Reading times at the most deeply embedded N") +
  theme_classic(base_size = 18) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 20,
      hjust = 1
    ),
    axis.line = element_line(linewidth = 1),
    axis.ticks = element_line(linewidth = 1)
  )
```

Here we measure specifically the effect which Kim et al. (2025) found significant: *who~obj~ and how* vs. *why* at the most deeply embedded N.

```{r, echo=TRUE}
# Omit that, just compare who_obj and how with why
library(lme4)
library(lmerTest)

data_no_that <- data_embedded %>%
  filter(whword != 'that') %>%
  mutate(is_why = ifelse((whword == 'why'), "why", "who_obj and how"))

model_2 <- lmer(
  log_rt ~ is_why + 
    (1 | participant_id) + 
    (1 | item_id),
  data = data_no_that
)

# View results
summary(model_2)
```

```{}
```

```{r, echo=TRUE}

```

Here the effect size points in the same direction as Kim et al. (2024) – *why* items are read faster than *who*/how*, but the effect is no longer significant.*

```{r, echo=TRUE}
# Omit why, just compare who_obj and how with why
data_no_that_no_how <- data_no_that %>%
  filter(whword != 'word') %>%
  mutate(is_why = ifelse((whword == 'why'), "why", "who_obj"))

model_3 <- lmer(
  log_rt ~ is_why + 
    (1 | participant_id) + 
    (1 | item_id),
  data = data_no_that_no_how
)

# View results
summary(model_3)
```

```{r}
agr <- data %>%
  group_by(word_index, whword) %>%
  summarise(y = mean(log_rt))
agr
```

```{r}
words <- c(
  "the",
  "father",
  "considered",
  "wh-element",
  "the",
  "naive",
  "babysitter",
  "that",
  "the",
  "spunky",
  "children",
  "loved",
  "unexpectedly",
  "and",
  "benevolently",
  "handed",
  "the",
  "toys",
  "to",
  "the",
  "grandma"
)

agr %>% 
  mutate(` ` = whword) %>%
  ggplot(aes(x = word_index, y = y,
             color = ` `,
             group = ` `)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  theme(
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 1
    )
  ) +
  scale_x_continuous(
    breaks = seq_along(words) - 1,
    labels = words
  ) +
  ggtitle("Means of region-by-region reading times") +
  labs(x="Words", y="Mean reading time (log ms)")
```
