---
title: "Power B analysis"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(pwr)        
library(emmeans)
theme_set(theme_minimal())


files <- c(
  c(
    "0j17p08k.csv", "1dg2qhj2b3.csv", "1e17ytgq.csv", "1fqrd6qr.csv", "1lmu0rmy.csv", "1orqqvyp.csv",
    "1qr6nnt1.csv", "1sftbbwm.csv", "2bomnjjb.csv", "2ea7uxyc.csv", "2ewef3gm.csv","2odgntx7.csv","3e3v39z8.csv","3jtr11xk.csv","3og97j2k.csv","3ut1qsdd.csv","4gjn1seg.csv","4qcemy4k.csv","4yxop4r3.csv","5lqc6nvcre.csv","5vrdpq0t.csv","7f2wtp1j.csv","7lhlpfus.csv","7mza4pym.csv","9kd8a7eb.csv","9nvsempv62.csv","20frs6ol.csv","31by1doq.csv","91mfkpae.csv","120ypmwg.csv","352jd0sb.csv","970gbe0a.csv","a7afpb9w.csv","adbunmf0.csv","ajl3gqra.csv","bn7akg4o.csv","bo3a253d.csv","bqlvmxzf.csv","c1m9yfql.csv","cktudb9s.csv","dsfc282a.csv","e34duc56.csv","el2hvftl.csv","gfzoatpu.csv","h8nlnku2.csv","h6253x9m.csv","hb3mv8ph.csv","hb4brkxq.csv","hs2xklzn.csv","hwsqhjot.csv","jrj7xzuf.csv","k2b5yzkl.csv","k8rnkquz.csv","kl8k8j1y.csv","lta397lq.csv","luzwjk24.csv","lwvpwcq0.csv","m1eabfx1.csv","n8adea4h.csv","na6hh0ua.csv","nflshhgq.csv","o92u56uh.csv","ogyjt8yg.csv","ojv4jxxa.csv","olol1tte.csv","orybxbxu.csv","p3taub1f.csv","pco9vqna.csv","pm5zycnt.csv","ppt63ukd.csv","q330tno2.csv","qbwjdear.csv","qfm8ks37.csv","qgzx1bfu.csv","r0pm70sc.csv","r1o8o6o4.csv","repvu995.csv","rxur173z.csv","s3j977x7.csv","szh4zjou.csv","tf6nb9bm.csv","tpt94g2b.csv","tw3k60gy.csv","txdtrb48.csv","u0pme1hk.csv","u1s85lkk.csv","u2fs8oqd.csv","ucyeq6mc.csv","uu3k1546.csv","vg74l3ea.csv","vgp86ou0.csv","vp4hn41a.csv","w3yqa2f3.csv","xthmgyn8.csv","y14x4t45.csv","yu3vjzn3.csv","z25ltc96.csv"
  )
)

# Read and combine all files from ../stims
data.raw <- files %>%
  set_names() %>%
  map_dfr(
    ~ suppressMessages(
        suppressWarnings(
          read_csv(
            file.path("../data/pilotb", .x),
            show_col_types = FALSE
          ) %>%
          mutate(rt = as.numeric(rt))
        )
      ),
    .id = "source_file"
  )
```

```{r}
# Load and preprocess dem data and stim data
dem_data <- read.csv("../data/pilot_b_demographic_data.csv")

# Clean up col names in dem_data
colnames(dem_data) <- c("submission_id", "participant_id", "status", "custom_study",
                        "started_at", "completed_at", "reviewed_at", "archived_at",
                        "time_taken", "completion_code", "total_approvals", "age",
                        "sex", "ethnicity", "birth_country", "residence_country",
                        "nationality", "language", "student", "employment")

dem_data <- dem_data %>%
  mutate(age = as.numeric(age))

stims.raw <- read.csv("../stims/stims_abridged_all_info.csv")
```

```{r}
# Remove filler and practice entries
# Focus on only target word for now (i.e. deepest NP in clause)
data <- data.raw %>%
  select(!stimulus) %>%
  left_join(dem_data %>% filter(status != "RETURNED") %>% select("participant_id", "age", "sex", "ethnicity", "birth_country",
                                "nationality", "language", "student", "employment"),
            by = join_by(subject_id == participant_id)) %>%
  left_join(stims.raw %>% select(item_id, target_word, stimulus), multiple = "first", by = join_by(item_id)) %>%
  mutate(rt = as.numeric(rt),
         log_rt = log(rt),
         whword = factor(ifelse(condition == 'a', 'object who', 
                         ifelse(condition == 'b', 'how',
                                ifelse(condition == 'c', 'that',
                                       'why'))), levels = c("object who", "how", "that", "why"))) %>% # condition = wh-word used in sentence
  filter(item_type != 'filler' & item_type != 'practice' & too_fast == FALSE) %>%
  filter(whword != "object who" | word_index < 19) %>%
  mutate(is_target_word = (target_word == word)) # select only deepest embedded NP
  

write.csv(data, "../data/pilot_b_data.csv")
```

Here we only include reading time observations taken at the most deeply embedded N. Further, we eliminate all observations where the reading time \< 100 ms (done via the condition too_fast == FALSE).

```{r}
data_embedded <- data %>% filter(is_target_word) # select only depeest embedded NP
```

Description of each field:

-   source_file: the original file in the directory 'data' storing a given participant's data

-   participant_id: 4-char participant id provided by Prolific

-   rt: reading time (log_rt: log transformed reading time)

-   condition: value from 'a'-'d' describing type of stimulus sentence

-   word: word at which reading time was measured

-   whword: one of four values "object who", "why", "how", or "that" – denotes the type of relative clause in the stimulus

```{r}
# Compute mean reaction time per condition
agr <- data_embedded %>%
  group_by(whword) %>%
  summarise(n = n(), se = sd(log_rt) / sqrt(n()), y = mean(log_rt)) %>%
  mutate(
    whword = factor(
      whword,
      levels = c("object who", "how", "that", "why")
    ) )# Order whword elements to be object who, how, that, why

agr

agr %>%
  mutate(` ` = whword) %>%
  ggplot(aes(x = ` `, y = y, fill = ` `)) +
  geom_bar(stat = "identity", width = 0.6, linewidth = 1) +
  geom_errorbar(aes(ymin = y - se, ymax = y + se ),   
                width = 0.08,   linewidth = 1 ) +
  scale_fill_manual(values = c("object who" = "blue",
                               "how" = "#D55E00",
                               "that" = "darkgreen",
                               "why" = "magenta")) +
  coord_cartesian(ylim = c(5, 5.6)) +
  labs(x = "wh-types", y = "Reaction time (log ms)") +
  ggtitle("Reading times at the most deeply embedded N") +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 20,
      hjust = 1
    ),
    axis.line = element_line(linewidth = 1),
    axis.ticks = element_line(linewidth = 1)
  ) +
  theme_minimal()
```

```{r}

library(lme4)
library(lmerTest)

data_embedded <- data_embedded %>%
  mutate(is_that = factor(ifelse((whword == 'that'), "that", "who_obj, how, or why"),
                levels = c("who_obj, how, or why", "that")))  

model_1 <- lmer(
  log_rt ~ is_that + 
    (1 | participant_id) + 
    (1 | item_id),
  data = data_embedded
)

# View results
summary(model_1)
```

Here we measure specifically the effect which Kim et al. (2025) found significant: *who~obj~ and how* vs. *why* at the most deeply embedded N.

```{r, echo=TRUE}
# Testing the who_obj + how vs. why effect (reported significant)

data_no_that <- data_embedded %>%
  filter(whword != 'that') %>%
  mutate(is_why = ifelse((whword == 'why'), "why", "who_obj and how"))

model_2 <- lmer(
  log_rt ~ is_why + 
    (1 | participant_id) + 
    (1 | item_id),
  data = data_no_that
)

# View results
summary(model_2)
```

```         
```

```{r, echo=TRUE}

```

Here the effect size points in the same direction as Kim et al. (2024) – *why* items are read faster than *who*/how*, but the effect is no longer significant.*

```{r}
# Omit why, just compare who_obj and how with why
data_no_that_no_how <- data_no_that %>%
  filter(whword != 'word') %>%
  mutate(is_why = factor(
    ifelse((whword == 'how'), "who", "how"),
  levels = c("who", "how")))

model_3 <- lmer(
  log_rt ~ is_why + 
    (1 | participant_id) + 
    (1 | item_id),
  data = data_no_that_no_how
)

# View results
summary(model_3)
```

```{r, echo=TRUE}
# Omit why, just compare who_obj and how with why
data_no_that_no_how <- data_no_that %>%
  filter(whword != 'word') %>%
  mutate(is_why = ifelse((whword == 'why'), "why", "who_obj"))

model_3 <- lmer(
  log_rt ~ is_why + 
    (1 | participant_id) + 
    (1 | item_id),
  data = data_no_that_no_how
)

# View results
summary(model_3)
```

```{r}
agr <- data %>%
  group_by(word_index, whword) %>%
  summarise(y = mean(log_rt))
words <- c(
  "the",
  "father",
  "considered",
  "wh-element",
  "the",
  "naive",
  "babysitter",
  "that",
  "the",
  "spunky",
  "children",
  "loved",
  "unexpectedly",
  "and",
  "benevolently",
  "handed",
  "the",
  "toys",
  "to",
  "the",
  "grandma"
)

agr %>% 
  mutate(` ` = whword) %>%
  ggplot(aes(x = word_index, y = y,
             color = ` `,
             group = ` `)) +
  geom_line(linewidth = 1, alpha = 0.7) +
  geom_point(size = 2, alpha = 0.7) +
  scale_color_manual(values = c("object who" = "blue",
                               "how" = "#D55E00",
                               "that" = "darkgreen",
                               "why" = "purple")) +
  theme(
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 1
    )
  ) +
  scale_x_continuous(
    breaks = seq_along(words) - 1,
    labels = words
  ) +
  ggtitle("Means of region-by-region reading times") +
  labs(x="Words", y="Mean reading time (log ms)")
```

Cross-cut data by age

```{r}
# Create dummy value (age < 35), where 35 is the median age

agr <- data %>%
  filter(!is.na(age)) %>%
  mutate(is_old = factor(ifelse(age < 35, "Younger than 35", "35 or older"),
                          levels = c("Younger than 35", "35 or older"))) %>%
  group_by(is_old, word_index, whword) %>%
  summarise(y = mean(log_rt))

agr %>% 
  mutate(` ` = whword) %>%
  ggplot(aes(x = word_index, y = y,
             color = ` `,
             group = ` `)) +
  geom_line(linewidth = 1, alpha = 0.7) +
  geom_point(size = 2, alpha = 0.7) +
  scale_color_manual(values = c("object who" = "blue",
                               "how" = "#D55E00",
                               "that" = "darkgreen",
                               "why" = "purple")) +
  theme(
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 1
    )
  ) +
  scale_x_continuous(
    breaks = seq_along(words) - 1,
    labels = words
  ) +
  ggtitle("Means of region-by-region reading times, split by age") +
  labs(x="Words", y="Mean reading time (log ms)") +
  facet_wrap(vars(is_old))
```
